---
title: "Initial R01 dataset and subsetting"
output: html_document
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r pressure, echo=FALSE}
library(knitr)
opts_knit$set(root.dir = 'C:/Users/David Vu/Documents/Work/Manuscripts/2019-DENV-Pf/DENV-malaria-R-files')
#Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
```
#This code was rebuilt based on ARK's GitHub code
```{r}
# packages -----------------------------------------------------------------
library(tidyr)
library(zoo)
library(lubridate)
library(utils)
```
#Download data from REDCap
```{r}
setwd("/Users/David Vu/Documents/Work/Manuscripts/2019-DENV-Pf/DENV-malaria-R-files") #need to run on console
#library(redcapAPI)
#library(REDCapR)
#Redcap.token <- readLines("C:/Users/David Vu/Documents/Work/REDCap Data/LaBeaud R01/RedCapToken.txt") # Read API token from folder
#REDcap.URL  <- 'https://redcap.stanford.edu/api/'
#rcon <- redcapConnection(url=REDcap.URL, token=Redcap.token)

#export data from redcap to R (must be connected via cisco VPN)
#R01_lab_results <- redcap_read(redcap_uri  = REDcap.URL, token= Redcap.token, batch_size = 300, continue_on_error = TRUE) #had technical problems downloading in chunks

#R01<-redcap_read_oneshot(redcap_uri= REDcap.URL, token=Redcap.token, raw_or_label = "raw", raw_or_label_headers = "raw", guess_type = FALSE, guess_max = 1000L, verbose = TRUE, config_options = NULL)$data #This worked
library(readr)
R01 <- read_csv("C:/Users/David Vu/Documents/Work/Manuscripts/2019-DENV-Pf/DENV-malaria-R-files/200407-R01.csv")
#load("2020-04-10_R01.rda") #most recent working file, 84922 obs 1597 variables
#R01backup<-R01

#currentDate <- Sys.Date() 
#FileName <- paste(Sys.Date(),"R01.rda", sep="_") 
#save(R01, file=FileName)
#write.csv(as.data.frame(R01), "200407-R01.csv", na="", row.names = F )
```
#AIC subset
```{r}
#subset AIC based on second character of person_id=="F"
table(substr(R01$person_id, 2, 2), exclude = NULL)#C:31069 F:50702 M:46 
#"M" due to person_ids "DMxxxx" from DIS study. These should correspond to existing AIC person_ids. Exclude from analysis.
#Need to merge data for these person_ids with original AIC
#Subset based on second character of person_id==F
AIC<-R01[which(substr(R01$person_id, 2, 2)=="F"),] #has 50800 obs
save(AIC, file=paste(Sys.Date(),"AIC.rda", sep="_"))
#-------------------------------
#Exclude visits not applicable to this analysis
AIC<- AIC[which(AIC$redcap_event_name!="visit_a2_arm_1"&AIC$redcap_event_name!="visit_b2_arm_1"&AIC$redcap_event_name!="visit_c2_arm_1"&AIC$redcap_event_name!="visit_d2_arm_1"&AIC$redcap_event_name!="visit_c2_arm_1"&AIC$redcap_event_name!="visit_u24_arm_1"&AIC$redcap_event_name!="visit_u24_fu_arm_1"), ]
#-------------------------------
#Number of assigned person_id = screened participants
library(dplyr) 
n_distinct(AIC$person_id) #8014 screened
#Number of enrolled = those with redcap_event_name=="patient_informatio_arm_1"
table(AIC$redcap_event_name, exclude = NULL)#7545 enrolled
```
#Additional variables: id_site (C,K,R,M,U), id_region (C,W), rural (1,0)
```{r}
#Define sites, city, rural ------------------------------------------------------
table(AIC$site, exclude = NULL) #site variable does not provide actual study site
#create id_site variable with the actual site from the first letter of person_id
AIC$id_site<-substr(AIC$person_id, 1, 1)
table(AIC$id_site, exclude = NULL)
#create id_region to distinguish coast and west
AIC$id_region<-NA
AIC <- within(AIC, id_region[AIC$id_site=="M"] <- "C") #Msambweni
AIC <- within(AIC, id_region[AIC$id_site=="U"] <- "C") #Ukunda
AIC <- within(AIC, id_region[AIC$id_site=="K"] <- "W") #Kisumu
AIC <- within(AIC, id_region[AIC$id_site=="C"] <- "W") #Chulaimbo
AIC <- within(AIC, id_region[AIC$id_site=="R"] <- "W") #Mbaka Oromo
table(AIC$id_site, AIC$id_region, exclude = NULL)
##rural
AIC$rural<-NA
AIC <- within(AIC, rural[AIC$id_site=="C"] <- 1)
AIC <- within(AIC, rural[AIC$id_site=="R"] <- 1)
AIC <- within(AIC, rural[AIC$id_site=="M"] <- 1)
AIC <- within(AIC, rural[AIC$id_site=="K"] <- 0)
AIC <- within(AIC, rural[AIC$id_site=="U"] <- 0)
table(AIC$id_site, AIC$rural, exclude = NULL)
```
#Work in progress ----------------------------------
#Dates: A lot of missing data among the various date variables. Strategy is first to identify valid date ranges and see if there is any discordance between the various date columns. Then if there is no evidence of discordance, to use data from the different columns to complement each other.
```{r}
library(tidyr)
grep("date", names(AIC), value=TRUE)
head(AIC$interview_date_aic)
table(AIC$interview_date_aic, exclude = NULL)
n_distinct(AIC$interview_date_aic)#1380
sum(is.na(AIC$interview_date_aic))
n_distinct(AIC$interview_date) #only 38 disctinct dates?
n_distinct(AIC$visit_date) #empty
n_distinct(AIC$pedsql_date_parent) #1195
n_distinct(AIC$date_collected) #907
# clean interview dates -----------------------------------------
#start with interview_date_aic since it has the most values
table(AIC$interview_date_aic, exclude = NULL) #12098 values of 1900-01-01 which should be NA
#create a new variable to use for visit date
table(AIC$visit_date, exclude = NULL) #existing visit_date is empty so will use that
#populate visit_date with interview_date_aic data
AIC$visit_date<-AIC$interview_date_aic
#change 1900-01-01 to NA
AIC$visit_date[AIC$visit_date=="1900-01-01"]<-NA
sum(is.na(AIC$interview_date_aic))#24382 NA's in original column
sum(is.na(AIC$visit_date))#36482 NA's in new column reflecting previous 12098 1900-01-01
table(AIC$visit_date, exclude = NULL) #no other values seem out of range
#try to fill NA from other date columns
table(AIC$interview_date[which(is.na(AIC$visit_date))], exclude = NULL) #42 values from interview_date
table(AIC$interview_date[which(is.na(AIC$visit_date))], AIC$pedsql_date_parent[which(is.na(AIC$visit_date))],exclude = NULL) #4 values match, but no discordance (other values complement each other); probably a more efficient way than manually examine but I don't know how
match(AIC$interview_date[which(is.na(AIC$visit_date))], AIC$pedsql_date_parent[which(is.na(AIC$visit_date))])
table(AIC$interview_date[which(is.na(AIC$visit_date))], AIC$date_collected[which(is.na(AIC$visit_date))],exclude = NULL)


AIC <- within(AIC, visit_date[AIC$id_site=="C"] <- 1)

R01_lab_results_visit<- R01_lab_results[which(R01_lab_results$redcap_event_name!="patient_informatio_arm_1"&R01_lab_results$redcap_event_name!="visit_u24_arm_1")  , ]
interview_dates<-R01_lab_results_visit[, grepl("person_id|redcap_event_name|interview_date|id_city|redcap_event_name|id_cohort", names(R01_lab_results_visit))]
interview_dates<-interview_dates[, !grepl("u24", names(interview_dates))]
interview_dates$interview_date_aic<-as.character(interview_dates$interview_date_aic)
interview_dates$interview_date_aic[is.na(interview_dates$interview_date_aic)] <- ""

interview_dates$interview_date<-as.character(interview_dates$interview_date)
interview_dates$interview_date[is.na(interview_dates$interview_date)] <- ""

interview_dates<-tidyr::unite(interview_dates, int_date, interview_date_aic:interview_date, sep='')
R01_lab_results<- merge(interview_dates, R01_lab_results,  by=c("person_id", "redcap_event_name", "id_city", "id_cohort"), all = TRUE)

class(R01_lab_results$int_date)

R01_lab_results$int_date <-lubridate::ymd(R01_lab_results$int_date)
dplyr::n_distinct(R01_lab_results$int_date)

R01_lab_results$month_year <- as.yearmon(R01_lab_results$int_date)

R01_lab_results$year <-lubridate::year(as.Date(R01_lab_results$int_date, origin = '1900-1-1'))

```
#Need to parse out symptoms because they were entered into a single column
```{r}
table(AIC$symptoms, exclude = NULL) #All NA, empty column
table(AIC$symptoms_aic, exclude = NULL) #symptoms listed in this variable
table(AIC$oth_symptoms_aic, exclude = NULL) #additional symptoms?
#create binary symptom vars -----------------------------------------------------------------
#subset symptoms
symptoms<-AIC[c("person_id", "redcap_event_name","symptoms", "symptoms_aic", "oth_symptoms_aic", "visit_type")]
symptomsback<-symptoms
symptoms<-symptomsback
n_distinct(symptoms$symptoms) #1 (all NA)
n_distinct(symptoms$symptoms_aic) #2481
n_distinct(symptoms$oth_symptoms_aic) #230, free text, first letter capitalized
#Amy's code: create binary symptom vars -----------------------------------------------------------------
#subset symptoms
#symptoms<-R01_lab_results[which(R01_lab_results$id_visit > 0), c("person_id", "redcap_event_name","symptoms", "symptoms_aic", "id_cohort", "id_city", "id_visit", "visit_type")]
#combine symptoms from different variables into a single variable
symptoms$all_symptoms<-NA
symptoms$all_symptoms<-symptoms$symptoms_aic
#symptoms$all_symptoms<-paste(symptoms$symptoms_aic, symptoms$oth_symptoms_aic, sep=" ")
symptoms<-as.data.frame(symptoms)
#all_symptoms has a lot of "NA" "Nil" and "None" from oth_symptoms_aic so need to remove those first: They are all capitalized
symptoms <- lapply(symptoms, function(x) {gsub("NA", "", x)})
grep("NA", symptoms$all_symptoms)
symptoms <- lapply(symptoms, function(x) {gsub("Nil", "", x)})
grep("Nil", symptoms$all_symptoms)
symptoms <- lapply(symptoms, function(x) {gsub("None", "", x)})
grep("None", symptoms$all_symptoms)
symptoms <- lapply(symptoms, function(x) {gsub("None", "", x)})
grep("No", symptoms$all_symptoms)
#make everything lower case for ease of replacement of all the rest
symptoms$all_symptoms<-tolower(symptoms$all_symptoms)
n_distinct(symptoms$all_symptoms) #2661 different unique all_symptoms strings

table(symptoms$oth_symptoms_aic)
#----------
symptoms$all_symptoms <- gsub("abdominal distension", 'ab_distend', symptoms$all_symptoms)
symptoms$all_symptoms <- gsub("abdominal distention", 'ab_distend', symptoms$all_symptoms)
symptoms$all_symptoms <- gsub("abdominal swelling", 'ab_distend', symptoms$all_symptoms)
symptoms$all_symptoms <- gsub("spleenomegaly", 'ab_distend', symptoms$all_symptoms)
#----------
symptoms$all_symptoms <- gsub("abdominal pain following injury 4/7 ago from a blunt object, child was hit and complains of pain", 'abp', symptoms$all_symptoms)
symptoms$all_symptoms <- gsub("lower abdominal pains", 'abp', symptoms$all_symptoms)
symptoms$all_symptoms <- gsub("lower abdominal pain", 'abp', symptoms$all_symptoms)
symptoms$all_symptoms <- gsub("abdominal pain", 'abp', symptoms$all_symptoms)
symptoms$all_symptoms <- gsub("epigastric pain", 'abp', symptoms$all_symptoms)
symptoms$all_symptoms <- gsub("epigastric", 'abp', symptoms$all_symptoms)
symptoms$all_symptoms <- gsub("epigastrin pain", 'abp', symptoms$all_symptoms)
symptoms$all_symptoms <- gsub("epigastrin", 'abp', symptoms$all_symptoms)
symptoms$all_symptoms <- gsub("epigast", 'abp', symptoms$all_symptoms)
symptoms$all_symptoms <- gsub("stomachache", 'abp', symptoms$all_symptoms)
symptoms$all_symptoms <- gsub("stomach", 'abp', symptoms$all_symptoms)

#grep("abdominal_pain", symptoms$all_symptoms)
#symptoms$all_symptoms <- gsub('abdomil_pain', 'abp', symptoms$all_symptoms)
#symptoms$all_symptoms <- gsub('abdominal_pa', 'abp', symptoms$all_symptoms)
#symptoms$all_symptoms <- gsub('abdomin', 'abp', symptoms$all_symptoms)
#symptoms$all_symptoms <- gsub('abdomina', 'abp', symptoms$all_symptoms)
#symptoms$all_symptoms <- gsub('abdominal_pai', 'abp', symptoms$all_symptoms)
#symptoms$all_symptoms <- gsub('abdominal_p', 'abp', symptoms$all_symptoms)
#symptoms$all_symptoms <- gsub('abdo', 'abp', symptoms$all_symptoms)
#symptoms$all_symptoms <- gsub('abd', 'abp', symptoms$all_symptoms)
#symptoms$all_symptoms <- gsub('abpin', 'abp', symptoms$all_symptoms)
#symptoms$all_symptoms <- gsub('abpi', 'abp', symptoms$all_symptoms)
#symptoms$all_symptoms <- gsub('abpal_p', 'abp', symptoms$all_symptoms)
#symptoms$all_symptoms <- gsub('abpa', 'abp', symptoms$all_symptoms)

symptoms$all_symptoms <- gsub('abp', 'abdominal_pain', symptoms$all_symptoms)
#----------
symptoms$all_symptoms <- gsub('exertional dyspnea', 'dyspnea', symptoms$all_symptoms)
symptoms$all_symptoms <- gsub('shortness_of_breath', 'dyspnea', symptoms$all_symptoms)
symptoms$all_symptoms <- gsub('difficulty in breathing', 'dyspnea', symptoms$all_symptoms)
symptoms$all_symptoms <- gsub('difficult in breathing', 'dyspnea', symptoms$all_symptoms)
symptoms$all_symptoms <- gsub('hard breathing', 'dyspnea', symptoms$all_symptoms)
symptoms$all_symptoms <- gsub('fast breathig', 'dyspnea', symptoms$all_symptoms)
symptoms$all_symptoms <- gsub('fast breathing', 'dyspnea', symptoms$all_symptoms)
symptoms$all_symptoms <- gsub("asthma", 'dyspnea', symptoms$all_symptoms)
symptoms$all_symptoms <- gsub("wheezing", 'dyspnea', symptoms$all_symptoms)
#----------         
symptoms$all_symptoms <- gsub("aneamia", 'pallor', symptoms$all_symptoms)
symptoms$all_symptoms <- gsub("looking pale", 'pallor', symptoms$all_symptoms)
symptoms$all_symptoms <- gsub("pale eyes", 'pallor', symptoms$all_symptoms)
symptoms$all_symptoms <- gsub("pale skin", 'pallor', symptoms$all_symptoms)
#----------
symptoms$all_symptoms <- gsub('loss of appetite', 'anorexia', symptoms$all_symptoms)
symptoms$all_symptoms <- gsub('loss_of_appetite', 'anorexia', symptoms$all_symptoms)
symptoms$all_symptoms <- gsub('refusal to feed', 'anorexia', symptoms$all_symptoms)
symptoms$all_symptoms <- gsub('lack of appetite', 'anorexia', symptoms$all_symptoms)
#---------- 
symptoms$all_symptoms <- gsub('chest cough', 'cough', symptoms$all_symptoms)
#----------
symptoms$all_symptoms <- gsub('convulsions', 'seizures', symptoms$all_symptoms)
symptoms$all_symptoms <- gsub('convulsion', 'seizures', symptoms$all_symptoms)
symptoms$all_symptoms <- gsub('conversion', 'seizures', symptoms$all_symptoms)
symptoms$all_symptoms <- gsub('fits', 'seizures', symptoms$all_symptoms)
#----------
symptoms$all_symptoms <- gsub('shivering', 'chills', symptoms$all_symptoms)
symptoms$all_symptoms <- gsub('freezing', 'chills', symptoms$all_symptoms)
symptoms$all_symptoms <- gsub('cold extremities', 'chills', symptoms$all_symptoms)
symptoms$all_symptoms <- gsub('chiils', 'chills', symptoms$all_symptoms)
#----------
symptoms$all_symptoms <- gsub('hard stool', 'constipation', symptoms$all_symptoms)
symptoms$all_symptoms <- gsub('failure to pass stool for 1 day', 'constipation', symptoms$all_symptoms)
symptoms$all_symptoms <- gsub('failure to pass stool', 'constipation', symptoms$all_symptoms)
symptoms$all_symptoms <- gsub('tenesmus', 'constipation', symptoms$all_symptoms)
#----------
symptoms$all_symptoms <- gsub('rti', 'dysuria', symptoms$all_symptoms)
symptoms$all_symptoms <- gsub('uti', 'dysuria', symptoms$all_symptoms)
symptoms$all_symptoms <- gsub('pain on urination', 'dysuria', symptoms$all_symptoms)
symptoms$all_symptoms <- gsub('pain while passing urine', 'dysuria', symptoms$all_symptoms)
symptoms$all_symptoms <- gsub('painful micturation', 'dysuria', symptoms$all_symptoms)
symptoms$all_symptoms <- gsub('painful urination', 'dysuria', symptoms$all_symptoms)
symptoms$all_symptoms <- gsub('frequency urination', 'dysuria', symptoms$all_symptoms)
#----------
symptoms$all_symptoms <- gsub('loose stool', 'diarrhea', symptoms$all_symptoms)
symptoms$all_symptoms <- gsub('foul smelly stool', 'diarrhea', symptoms$all_symptoms)
#----------
symptoms$all_symptoms <- gsub('itchyness', 'itchiness', symptoms$all_symptoms)
symptoms$all_symptoms <- gsub('skin itching', 'itchiness', symptoms$all_symptoms)
symptoms$all_symptoms <- gsub('itching at the penile shaft', 'itchiness', symptoms$all_symptoms)
symptoms$all_symptoms <- gsub('longstanding itchy and painful eyes', 'itchiness', symptoms$all_symptoms)
symptoms$all_symptoms <- gsub('itchy eyes', 'itchiness', symptoms$all_symptoms)
symptoms$all_symptoms <- gsub('eye itching', 'itchiness', symptoms$all_symptoms)
#----------
symptoms$all_symptoms <- gsub('hotness of the body', 'fever', symptoms$all_symptoms)
symptoms$all_symptoms <- gsub('febrile', 'fever', symptoms$all_symptoms)
#----------
symptoms$all_symptoms <- gsub('running nose', 'runny_nose', symptoms$all_symptoms)
symptoms$all_symptoms <- gsub('runing nose', 'runny_nose', symptoms$all_symptoms)
#----------
symptoms$all_symptoms <- gsub('joint pains', 'joint_pains', symptoms$all_symptoms)
symptoms$all_symptoms <- gsub('hip pain', 'joint_pains', symptoms$all_symptoms)
#----------
symptoms$all_symptoms <- gsub("back ache", 'back_pain', symptoms$all_symptoms)
symptoms$all_symptoms <- gsub("lower back pains", 'back_pain', symptoms$all_symptoms)
symptoms$all_symptoms <- gsub("left frankpain", 'back_pain', symptoms$all_symptoms)
symptoms$all_symptoms <- gsub("back pain", 'back_pain', symptoms$all_symptoms)
#----------
symptoms$all_symptoms <- gsub('foul smell from the mouth', 'halitosis', symptoms$all_symptoms)
symptoms$all_symptoms <- gsub('mouth odour', 'halitosis', symptoms$all_symptoms)
symptoms$all_symptoms <- gsub('strong foul smell from the mouth', 'halitosis', symptoms$all_symptoms)
#----------
symptoms$all_symptoms <- gsub('lethergy', 'lethargy', symptoms$all_symptoms)
#----------
symptoms$all_symptoms <- gsub('chest pain', 'chest_pain', symptoms$all_symptoms)
symptoms$all_symptoms <- gsub('chestpain', 'chest_pain', symptoms$all_symptoms)
#----------
symptoms$all_symptoms <- gsub('ulceration on the tongue', 'mouth_sores', symptoms$all_symptoms)
symptoms$all_symptoms <- gsub('wounds in the mouth', 'mouth_sores', symptoms$all_symptoms)
symptoms$all_symptoms <- gsub('oral lesion', 'mouth_sores', symptoms$all_symptoms)
symptoms$all_symptoms <- gsub('oral sores', 'mouth_sores', symptoms$all_symptoms)
symptoms$all_symptoms <- gsub('oral sore', 'mouth_sores', symptoms$all_symptoms)
symptoms$all_symptoms <- gsub('mouth sores', 'mouth_sores', symptoms$all_symptoms)
symptoms$all_symptoms <- gsub('mouth sore', 'mouth_sores', symptoms$all_symptoms)
#----------
symptoms$all_symptoms <- gsub('rash in the head', 'rash', symptoms$all_symptoms)
symptoms$all_symptoms <- gsub('rashes on face', 'rash', symptoms$all_symptoms)
symptoms$all_symptoms <- gsub('post inflammation Skin lesions', 'rash', symptoms$all_symptoms)
symptoms$all_symptoms <- gsub('skin blister', 'rash', symptoms$all_symptoms)
symptoms$all_symptoms <- gsub('septic rash', 'rash', symptoms$all_symptoms)
symptoms$all_symptoms <- gsub('skin rash', 'rash', symptoms$all_symptoms)
symptoms$all_symptoms <- gsub('sores on the neck', 'rash', symptoms$all_symptoms)
symptoms$all_symptoms <- gsub('sun rash', 'rash', symptoms$all_symptoms)
symptoms$all_symptoms <- gsub('rashes', 'rash', symptoms$all_symptoms)
#----------
symptoms$all_symptoms <- gsub('sore throats', 'sore_throat', symptoms$all_symptoms)
symptoms$all_symptoms <- gsub('sore throat', 'sore_throat', symptoms$all_symptoms)
symptoms$all_symptoms <- gsub('inflamed tonsils', 'sore_throat', symptoms$all_symptoms)
symptoms$all_symptoms <- gsub('tonsolitis', 'sore_throat', symptoms$all_symptoms)
symptoms$all_symptoms <- gsub('tonsilitis', 'sore_throat', symptoms$all_symptoms)
symptoms$all_symptoms <- gsub('tonsillitis', 'sore_throat', symptoms$all_symptoms)
symptoms$all_symptoms <- gsub('tonsils', 'sore_throat', symptoms$all_symptoms)
symptoms$all_symptoms <- gsub('pharyngititis', 'sore_throat', symptoms$all_symptoms)
symptoms$all_symptoms <- gsub('pain when swallowing', 'sore_throat', symptoms$all_symptoms)
symptoms$all_symptoms <- gsub('ain on swallowing', 'sore_throat', symptoms$all_symptoms)
symptoms$all_symptoms <- gsub('odinophagia', 'sore_throat', symptoms$all_symptoms)
symptoms$all_symptoms <- gsub('dysphagia', 'sore_throat', symptoms$all_symptoms)
symptoms$all_symptoms <- gsub('throat', 'sore_throat', symptoms$all_symptoms)
#----------
symptoms$all_symptoms <- gsub('fainting', 'altered_ms', symptoms$all_symptoms)
symptoms$all_symptoms <- gsub('hallucination', 'altered_ms', symptoms$all_symptoms)
#----------
symptoms$all_symptoms <- gsub('watery eye discharge', 'eye_discharge', symptoms$all_symptoms)
symptoms$all_symptoms <- gsub('whitish eye discharge', 'eye_discharge', symptoms$all_symptoms)
symptoms$all_symptoms <- gsub('tearing eyes', 'eye_discharge', symptoms$all_symptoms)
symptoms$all_symptoms <- gsub('eye. discharge', 'eye_discharge', symptoms$all_symptoms)
symptoms$all_symptoms <- gsub('eye discharge', 'eye_discharge', symptoms$all_symptoms)
symptoms$all_symptoms <- gsub('discharging eyes', 'eye_discharge', symptoms$all_symptoms)
symptoms$all_symptoms <- gsub('eye infection', 'eye_discharge', symptoms$all_symptoms)
#----------
symptoms$all_symptoms <- gsub('pus ear discharge', 'ear_discharge', symptoms$all_symptoms)
symptoms$all_symptoms <- gsub('ear discharge', 'ear_discharge', symptoms$all_symptoms)
#----------
symptoms$all_symptoms <- gsub('irritability', 'irritable', symptoms$all_symptoms)
symptoms$all_symptoms <- gsub('restlessness', 'irritable', symptoms$all_symptoms)
symptoms$all_symptoms <- gsub('restless', 'irritable', symptoms$all_symptoms)
#----------
symptoms$all_symptoms <- gsub('bilateral swelling of the chin', 'facial_swelling', symptoms$all_symptoms)
symptoms$all_symptoms <- gsub('swollen face', 'facial_swelling', symptoms$all_symptoms)
symptoms$all_symptoms <- gsub('Swelling on the right chin', 'facial_swelling', symptoms$all_symptoms)
symptoms$all_symptoms <- gsub('swollen cheeks', 'facial_swelling', symptoms$all_symptoms)
symptoms$all_symptoms <- gsub('swollen eyes', 'facial_swelling', symptoms$all_symptoms)
symptoms$all_symptoms <- gsub('puffiness of the face', 'facial_swelling', symptoms$all_symptoms)
symptoms$all_symptoms <- gsub('puffy eyes', 'facial_swelling', symptoms$all_symptoms)
symptoms$all_symptoms <- gsub('infra auricular sweling', 'facial_swelling', symptoms$all_symptoms)
symptoms$all_symptoms <- gsub('intraorbital swellin', 'facial_swelling', symptoms$all_symptoms)
symptoms$all_symptoms <- gsub('swelling below the left ear', 'facial_swelling', symptoms$all_symptoms)
symptoms$all_symptoms <- gsub('swelling on the left ear', 'facial_swelling', symptoms$all_symptoms)
symptoms$all_symptoms <- gsub('eye swelling', 'facial_swelling', symptoms$all_symptoms)
#----------
symptoms$all_symptoms <- gsub('cervical lymph nodes', 'swollen_nodes', symptoms$all_symptoms)
symptoms$all_symptoms <- gsub('neckswelling', 'swollen_nodes', symptoms$all_symptoms)
symptoms$all_symptoms <- gsub('swelling on the neck', 'swollen_nodes', symptoms$all_symptoms)
symptoms$all_symptoms <- gsub('swollen inguinal lymhnodes', 'swollen_nodes', symptoms$all_symptoms)
#----------
symptoms$all_symptoms <- gsub('occasional sturtling', 'easily_startled', symptoms$all_symptoms)
symptoms$all_symptoms <- gsub('strutles', 'easily_startled', symptoms$all_symptoms)
symptoms$all_symptoms <- gsub('sturtling', 'easily_startled', symptoms$all_symptoms)
#----------
symptoms$all_symptoms <- gsub('bruises inside the nose', 'bruising', symptoms$all_symptoms)
#----------
symptoms$all_symptoms <- gsub('night sweat', 'night_sweats', symptoms$all_symptoms)
#----------
symptoms$all_symptoms <- gsub('Yellowness of the eyes', 'jaundice', symptoms$all_symptoms) 
#----------
symptoms$all_symptoms <- gsub('red eyes', 'red_eyes', symptoms$all_symptoms)
#----------
symptoms$all_symptoms <- gsub('boggy swelling of rt foot', 'edema', symptoms$all_symptoms)
#----------
symptoms$all_symptoms <- gsub('neck pain', 'myalgia', symptoms$all_symptoms)
symptoms$all_symptoms <- gsub('swelling on the right thigh', 'myalgia', symptoms$all_symptoms)
symptoms$all_symptoms <- gsub('swelling on the rt gluteas muscle', 'myalgia', symptoms$all_symptoms)
symptoms$all_symptoms <- gsub('muscle_pains', 'myalgia', symptoms$all_symptoms)
symptoms$all_symptoms <- gsub('general body pains', 'myalgia', symptoms$all_symptoms)

#----------
symptoms$all_symptoms <- gsub('tinea corporis', 'other', symptoms$all_symptoms)
symptoms$all_symptoms <- gsub('fungal skin infection', 'other', symptoms$all_symptoms) #Tinea corporis
symptoms$all_symptoms <- gsub('t.carpitis', 'other', symptoms$all_symptoms)#Tinea capitis
symptoms$all_symptoms <- gsub('wounds on the head', 'other', symptoms$all_symptoms) #Tinea capitis
symptoms$all_symptoms <- gsub('tooth ache', 'other', symptoms$all_symptoms)
symptoms$all_symptoms <- gsub('toothache and swelling on the gum', 'other', symptoms$all_symptoms)
symptoms$all_symptoms <- gsub('worms in his stool ', 'other', symptoms$all_symptoms)
symptoms$all_symptoms <- gsub('worms', 'other', symptoms$all_symptoms)
symptoms$all_symptoms <- gsub('chicken pox', 'other', symptoms$all_symptoms)
symptoms$all_symptoms <- gsub('ringworms', 'other', symptoms$all_symptoms)
symptoms$all_symptoms <- gsub('ringworm', 'other', symptoms$all_symptoms)
symptoms$all_symptoms <- gsub('schistomiasis', 'other', symptoms$all_symptoms)
symptoms$all_symptoms <- gsub('cravy for soil', 'other', symptoms$all_symptoms)
symptoms$all_symptoms <- gsub('jiggers', 'other', symptoms$all_symptoms) #chiggers
symptoms$all_symptoms <- gsub('oral thrush', 'other', symptoms$all_symptoms)
symptoms$all_symptoms <- gsub('sprained wrist', 'other', symptoms$all_symptoms)
symptoms$all_symptoms <- gsub('mumps', 'other', symptoms$all_symptoms)
symptoms$all_symptoms <- gsub('sunisitis', 'other', symptoms$all_symptoms)
symptoms$all_symptoms <- gsub('neutrophilis', 'other', symptoms$all_symptoms)
symptoms$all_symptoms <- gsub('nasal lacerations', 'other', symptoms$all_symptoms)
symptoms$all_symptoms <- gsub('dullness', 'other', symptoms$all_symptoms)
symptoms$all_symptoms <- gsub('fast paced heart beat', 'other', symptoms$all_symptoms)
symptoms$all_symptoms <- gsub('jaw pain', 'other', symptoms$all_symptoms)
symptoms$all_symptoms <- gsub('yellow urine', 'other', symptoms$all_symptoms)
symptoms$all_symptoms <- gsub('eye pain', 'other', symptoms$all_symptoms)
symptoms$all_symptoms <- gsub('swelling on the left thumb', 'other', symptoms$all_symptoms)
symptoms$all_symptoms <- gsub('swollen thumb', 'other', symptoms$all_symptoms)
#----------
symptoms$all_symptoms <- gsub('piriton', '', symptoms$all_symptoms) #allergy med not symptom
symptoms$all_symptoms <- gsub('amoxicillin', '', symptoms$all_symptoms) #med not symptom
#----------
symptoms$all_symptoms <- gsub('easy fatigue', 'fatigue', symptoms$all_symptoms)
symptoms$all_symptoms <- gsub('refusal to play', 'fatigue', symptoms$all_symptoms)

#----------
symptoms$all_symptoms <- gsub('sleeping alot', 'decreased_activity', symptoms$all_symptoms)
#----------
symptoms$all_symptoms <- gsub('ear pain', 'otalgia', symptoms$all_symptoms)
symptoms$all_symptoms <- gsub('earache', 'otalgia', symptoms$all_symptoms)
#symptoms$all_symptoms <- gsub('ear', 'otalgia', symptoms$all_symptoms)

symptoms$all_symptoms <- gsub('flew', 'flu', symptoms$all_symptoms)
symptoms$all_symptoms <- gsub('flue', 'flu', symptoms$all_symptoms)
#symptoms$all_symptoms <- gsub('cold', 'flu', symptoms$all_symptoms)
#----------
symptoms$all_symptoms <- gsub('general body malaise', 'sick_feeling', symptoms$all_symptoms)
symptoms$all_symptoms <- gsub('general malaise', 'sick_feeling', symptoms$all_symptoms)
symptoms$all_symptoms <- gsub('sick_feeling', 'malaise', symptoms$all_symptoms)
sympback<-symptoms
#----------
symptoms$all_symptoms <- gsub('septic wounds on the toes', 'skin_infection', symptoms$all_symptoms)
symptoms$all_symptoms <- gsub('boggy pus discharging swelling on buttock', 'skin_infection', symptoms$all_symptoms)
#symptoms$all_symptoms <- gsub('sores', 'skin_infection', symptoms$all_symptoms)


#symptoms$all_symptoms <- gsub('diarrh', 'diarrhea', symptoms$all_symptoms)
#symptoms$all_symptoms <- gsub('diarrheaea', 'diarrhea', symptoms$all_symptoms)
#symptoms$all_symptoms <- gsub('usea', 'nausea', symptoms$all_symptoms)
#symptoms$all_symptoms <- gsub('nanausea', 'nausea', symptoms$all_symptoms)
#symptoms$all_symptoms <- gsub('ras', 'rash', symptoms$all_symptoms)
#symptoms$all_symptoms <- gsub('rashh', 'rash', symptoms$all_symptoms)

#symptoms$all_symptoms <- gsub('99', '', symptoms$all_symptoms)
#symptoms$all_symptoms <- gsub('none', '', symptoms$all_symptoms)
#symptoms$all_symptoms <- gsub('no', '', symptoms$all_symptoms)
#symptoms$all_symptoms <- gsub('nil', '', symptoms$all_symptoms)
#symptoms$all_symptoms <- gsub("\\<a\\>", '', symptoms$all_symptoms)
grep(",", symptoms$all_symptoms)
#clean up commas and replace with space
symptoms <- lapply(symptoms, function(x) {gsub(",", " ", x)})
grep(",", symptoms$all_symptoms)
#clean up "na"
#symptoms <- lapply(symptoms, function(x) {gsub(",NA", "", x)})
#symptoms <- lapply(symptoms, function(x) {gsub("na", "", x)})
#grep("na", symptoms$all_symptoms)
#clean up "nil"
#symptoms <- lapply(symptoms, function(x) {gsub("nil", "", x)})
#grep("nil", symptoms$all_symptoms)
#clean up "none"
#symptoms <- lapply(symptoms, function(x) {gsub("none", "", x)})
#grep("none", symptoms$all_symptoms)
#symptoms <- lapply(symptoms, function(x) {gsub(",none", "", x)})
symptoms<-as.data.frame(symptoms)
sympback<-symptoms
#symptoms <-symptoms[!(is.na(symptoms$all_symptoms) | symptoms$all_symptoms==" "), ]
lev <- levels(factor(symptoms$all_symptoms))
lev <- unique(unlist(strsplit(lev, " ")))
mnames <- gsub(" ", "_", paste("sym", lev, sep = "_"))
result <- matrix(data = "0", nrow = length(symptoms$all_symptoms), ncol = length(lev))
char.aic_symptom <- as.character(symptoms$all_symptoms)
for (i in 1:length(lev)) {
  result[grep(lev[i], char.aic_symptom, fixed = TRUE), i] <- "1"
}
result <- data.frame(result, stringsAsFactors = TRUE)
colnames(result) <- mnames
symptoms <- cbind(symptoms,result)

as.numeric.factor <- function(x) {as.numeric(levels(x))[x]}
ids <- c("person_id", "redcap_event_name", "all_symptoms")
ids <- symptoms[ids]
symptoms<-symptoms[ , !(names(symptoms) %in% "sym_")]
aic_symptom_<-grep("sym_", names(symptoms), value = TRUE)
symptoms<-symptoms[ , (names(symptoms) %in% aic_symptom_)]
symptoms <-as.data.frame(sapply(symptoms, as.numeric.factor))

#collapse the symptoms that are the same.
#symptoms <- within(symptoms, aic_symptom_impaired_mental_status[symptoms$aic_symptom_fits==0|symptoms$aic_symptom_seizures==0] <- 0)
#symptoms <- within(symptoms, aic_symptom_impaired_mental_status[symptoms$aic_symptom_fits==1|symptoms$aic_symptom_seizures==1] <- 1)

symptoms$bleeding<-NA
symptoms <- within(symptoms, bleeding[symptoms$aic_symptom_bleeding_gums==0] <- 0)
symptoms <- within(symptoms, bleeding[symptoms$aic_symptom_bleeding_gums==0] <- 0)
symptoms <- within(symptoms, bleeding[symptoms$aic_symptom_bloody_nose==0] <- 0)
symptoms <- within(symptoms, bleeding[symptoms$aic_symptom_bloody_urine==0] <- 0)
symptoms <- within(symptoms, bleeding[symptoms$aic_symptom_bloody_stool==0] <- 0)
symptoms <- within(symptoms, bleeding[symptoms$aic_symptom_bloody_vomit==0] <- 0)
symptoms <- within(symptoms, bleeding[symptoms$aic_symptom_bruises==0] <- 0)

symptoms <- within(symptoms, bleeding[symptoms$aic_symptom_bleeding_gums==1] <- 1)
symptoms <- within(symptoms, bleeding[symptoms$aic_symptom_bleeding_gums==1] <- 1)
symptoms <- within(symptoms, bleeding[symptoms$aic_symptom_bloody_nose==1] <- 1)
symptoms <- within(symptoms, bleeding[symptoms$aic_symptom_bloody_urine==1] <- 1)
symptoms <- within(symptoms, bleeding[symptoms$aic_symptom_bloody_stool==1] <- 1)
symptoms <- within(symptoms, bleeding[symptoms$aic_symptom_bloody_vomit==1] <- 1)
symptoms <- within(symptoms, bleeding[symptoms$aic_symptom_bruises==1] <- 1)

symptoms$body_ache<-NA
symptoms <- within(symptoms, body_ache[symptoms$aic_symptom_general_body_ache==0] <- 0)
symptoms <- within(symptoms, body_ache[symptoms$aic_symptom_muscle_pains==0] <- 0)
symptoms <- within(symptoms, body_ache[symptoms$aic_symptom_bone_pains==0] <- 0)

symptoms <- within(symptoms, body_ache[symptoms$aic_symptom_general_body_ache==1] <- 1)
symptoms <- within(symptoms, body_ache[symptoms$aic_symptom_muscle_pains==1] <- 1)
symptoms <- within(symptoms, body_ache[symptoms$aic_symptom_bone_pains==1] <- 1)
table(symptoms$body_ache)
variable.names(symptoms)

#symptoms$nausea_vomitting<-NA
#symptoms <- within(symptoms, nausea_vomitting[symptoms$aic_symptom_nausea==0|symptoms$aic_symptom_vomiting==0| symptoms$aic_symptom_bloody_vomit==0] <- 0)
#symptoms <- within(symptoms, nausea_vomitting[symptoms$aic_symptom_nausea==1|symptoms$aic_symptom_vomiting==1| symptoms$aic_symptom_bloody_vomit==1] <- 1)

symptoms<-symptoms[ , grepl( "aic_symptom|bleeding|body_ache|nausea_vomitting" , names(symptoms) ) ]
symptoms$symptom_sum <- as.integer(rowSums(symptoms[ , grep("aic_symptom" , names(symptoms))], na.rm = TRUE))
table(symptoms$symptom_sum, exclude=NULL)
symptoms$symptomatic<-NA
symptoms <- within(symptoms, symptomatic[symptoms$symptom_sum>0] <- 1)
symptoms <- within(symptoms, symptomatic[symptoms$symptom_sum==0] <- 0)

#how much of our acute DENV was symptomatic vs. mildly/asymptomatic, etc.
table(symptoms$symptomatic, exclude=NULL)
#export to box.
symptoms<-as.data.frame(cbind(ids, symptoms))
symptoms$id_cohort<-substr(symptoms$person_id, 2, 2)
aic_symptoms<-subset(symptoms, id_cohort!="C")
aic_symptoms<-aic_symptoms[, !grepl("id_cohort", names(aic_symptoms))]

# create binary physical vars -----------------------------------------------------------------
#parce the physical exam results-----------------------------------------------------------------
#subset physical_exam
physical_exam<-R01_lab_results[which(R01_lab_results$redcap_event_name!="patient_informatio_arm_1"), c("person_id", "redcap_event_name","head_neck_exam", "skin", "neuro", "abdomen", "chest", "heart", "nodes", "joints")]
physical_exam<-as.data.frame(physical_exam)    
physical_exam$all_exam<-paste(physical_exam$head_neck_exam, physical_exam$skin, physical_exam$neuro, physical_exam$abdomen, physical_exam$joints, physical_exam$nodes, physical_exam$heart, physical_exam$chest, physical_exam$abdomen, sep=" ")
physical_exam <- lapply(physical_exam, function(x) {
  gsub(",NA", "", x)
})
physical_exam <- lapply(physical_exam, function(x) {
  gsub("NA", "", x)
})
physical_exam <- lapply(physical_exam, function(x) {
  gsub(",none", "", x)
})
physical_exam <- lapply(physical_exam, function(x) {
  gsub("none", "", x)
})
physical_exam<-as.data.frame(physical_exam)

#create dummy vars for all physical_exam
lev <- levels(factor(physical_exam$all_exam))
lev <- unique(unlist(strsplit(lev, " ")))
mnames <- gsub(" ", "_", paste("aic_pe", lev, sep = "_"))
result <- matrix(data = "0", nrow = length(physical_exam$all_exam), ncol = length(lev))
char.aic_pe <- as.character(physical_exam$all_exam)
for (i in 1:length(lev)) {
  result[grep(lev[i], char.aic_pe, fixed = TRUE), i] <- "1"
}
result <- data.frame(result, stringsAsFactors = TRUE)
colnames(result) <- mnames
physical_exam <- cbind(physical_exam,result)

as.numeric.factor <- function(x) {as.numeric(levels(x))[x]}
ids <- c("person_id", "redcap_event_name", "all_exam")
ids <- physical_exam[ids]
physical_exam<-physical_exam[ , !(names(physical_exam) %in% "aic_pe_")]
aic_pe<-grep("aic_pe_", names(physical_exam), value = TRUE)
physical_exam<-physical_exam[ , (names(physical_exam) %in% aic_pe)]
physical_exam <-as.data.frame(sapply(physical_exam, as.numeric.factor))

#create binary var for nodes      
physical_exam$node<-NA
table(physical_exam$aic_pe_large_lymph_nodes)

physical_exam<-as.data.frame(cbind(ids, physical_exam))

# reshape testing vars -----------------------------------------------------------------
#tested
tested<-R01_lab_results[, grepl("person_id|redcap_event|tested_", names(R01_lab_results))]
tested<-tested[, !grepl("date|freezer|rack|sample", names(tested))]

tested<-tested[which(tested$redcap_event_name=="patient_informatio_arm_1"),]
colnames(tested)<-sub("tested_*","", colnames(tested))    
colnames(tested)<-sub("*_2","", colnames(tested))    
#order
tested<-tested[,order(colnames(tested))]

tested<-tested[order(-(grepl('gh_$', names(tested)))+1L)]
tested<-tested[order(-(grepl('fg_$', names(tested)))+1L)]
tested<-tested[order(-(grepl('ef_$', names(tested)))+1L)]
tested<-tested[order(-(grepl('de_$', names(tested)))+1L)]
tested<-tested[order(-(grepl('cd_$', names(tested)))+1L)]
tested<-tested[order(-(grepl('bc_$', names(tested)))+1L)]
tested<-tested[order(-(grepl('ab_$', names(tested)))+1L)]

nameVec <- names(tested)
v.names=c('chikv_kenya_igg', 'chikv_stfd_igg', 'denv_kenya_igg', 'denv_stfd_igg')
times = c("ab_", "bc_", "cd_", "de_", "ef_", "fg_", "gh_")    
tested_long<-reshape(tested, idvar = "person_id", varying = 1:28,  direction = "long", timevar = "visit", times=times, v.names=v.names)
table(tested$ab_chikv_stfd_igg)
table(tested$bc_chikv_stfd_igg)
table(tested_long$chikv_stfd_igg, tested_long$visit)


tested_long <- within(tested_long, visit[visit=="ab_"] <- "visit_a_arm_1")
tested_long <- within(tested_long, visit[visit=="bc_"] <- "visit_b_arm_1")
tested_long <- within(tested_long, visit[visit=="cd_"] <- "visit_c_arm_1")
tested_long <- within(tested_long, visit[visit=="de_"] <- "visit_d_arm_1")
tested_long <- within(tested_long, visit[visit=="ef_"] <- "visit_e_arm_1")
tested_long <- within(tested_long, visit[visit=="fg_"] <- "visit_f_arm_1")
tested_long <- within(tested_long, visit[visit=="gh_"] <- "visit_g_arm_1")
tested_long$redcap_event_name<-tested_long$visit
names(tested_long)[names(tested_long) == 'denv_kenya_igg'] <- 'tested_denv_kenya_igg'
names(tested_long)[names(tested_long) == 'chikv_kenya_igg'] <- 'tested_chikv_kenya_igg'
names(tested_long)[names(tested_long) == 'denv_stfd_igg'] <- 'tested_denv_stfd_igg'
names(tested_long)[names(tested_long) == 'chikv_stfd_igg'] <- 'tested_chikv_stfd_igg'

head(tested_long)
# reshape sereoconverter vars -----------------------------------------------------------------
#seroconversion
seroconverter<-R01_lab_results[, grepl("person_id|redcap_event|ab_|bc_|cd_|de_|ef_|fg_|gh_", names(R01_lab_results))]
seroconverter<-seroconverter[, !grepl("malaria|tested|freezer|rack|sample|zcd", names(seroconverter))]
seroconverter<-seroconverter[which(seroconverter$redcap_event_name=="patient_informatio_arm_1"),]

#order
seroconverter<-seroconverter[,order(colnames(seroconverter))]

seroconverter<-seroconverter[order(-(grepl('gh_$', names(seroconverter)))+1L)]
seroconverter<-seroconverter[order(-(grepl('fg_$', names(seroconverter)))+1L)]
seroconverter<-seroconverter[order(-(grepl('ef_$', names(seroconverter)))+1L)]
seroconverter<-seroconverter[order(-(grepl('de_$', names(seroconverter)))+1L)]
seroconverter<-seroconverter[order(-(grepl('cd_$', names(seroconverter)))+1L)]
seroconverter<-seroconverter[order(-(grepl('bc_$', names(seroconverter)))+1L)]
seroconverter<-seroconverter[order(-(grepl('ab_$', names(seroconverter)))+1L)]

nameVec <- names(seroconverter)
v.names=c('chikv_kenya_igg', 'chikv_stfd_igg', 'denv_kenya_igg', 'denv_stfd_igg')
times = c("ab_", "bc_", "cd_", "de_", "ef_", "fg_", "gh_")    
seroconverter_long<-reshape(seroconverter, idvar = "person_id", varying = 1:28,  direction = "long", timevar = "visit", times=times, v.names=v.names)

table(seroconverter$ab_chikv_stfd_igg)
table(seroconverter$ab_denv_stfd_igg)
table(seroconverter$bc_chikv_stfd_igg)
table(seroconverter$bc_denv_stfd_igg)
table(seroconverter_long$chikv_stfd_igg, seroconverter_long$visit)
table(seroconverter_long$denv_stfd_igg, seroconverter_long$visit)

seroconverter_long <- within(seroconverter_long, visit[visit=="ab_"] <- "visit_a_arm_1")
seroconverter_long <- within(seroconverter_long, visit[visit=="bc_"] <- "visit_b_arm_1")
seroconverter_long <- within(seroconverter_long, visit[visit=="cd_"] <- "visit_c_arm_1")
seroconverter_long <- within(seroconverter_long, visit[visit=="de_"] <- "visit_d_arm_1")
seroconverter_long <- within(seroconverter_long, visit[visit=="ef_"] <- "visit_e_arm_1")
seroconverter_long <- within(seroconverter_long, visit[visit=="fg_"] <- "visit_f_arm_1")
seroconverter_long <- within(seroconverter_long, visit[visit=="gh_"] <- "visit_g_arm_1")
seroconverter_long$redcap_event_name<-seroconverter_long$visit
names(seroconverter_long)[names(seroconverter_long) == 'denv_kenya_igg'] <- 'seroc_denv_kenya_igg'
names(seroconverter_long)[names(seroconverter_long) == 'chikv_kenya_igg'] <- 'seroc_chikv_kenya_igg'
names(seroconverter_long)[names(seroconverter_long) == 'denv_stfd_igg'] <- 'seroc_denv_stfd_igg'
names(seroconverter_long)[names(seroconverter_long) == 'chikv_stfd_igg'] <- 'seroc_chikv_stfd_igg'

save(aic_symptoms,file="aic_symptoms.rda")
# merging the created data sets back to main -----------------------------------------------------------------
R01_lab_results <- merge(seroconverter_long, R01_lab_results,  by=c("person_id", "redcap_event_name"), all = TRUE)  #merge symptoms to redcap data
R01_lab_results <- merge(aic_symptoms, R01_lab_results, by=c("person_id", "redcap_event_name"), all = TRUE)  #merge symptoms to redcap data
R01_lab_results <- merge(physical_exam, R01_lab_results,  by=c("person_id", "redcap_event_name"), all = TRUE)  #merge pe parsed data
R01_lab_results <- merge(tested_long, R01_lab_results,  by=c("person_id", "redcap_event_name"), all = TRUE)  #merge tested samples 
# dengue : probable and warning -----------------------------------------------------------------
R01_lab_results$probable_dengue<-rowSums(R01_lab_results[, grep("/body_ache|aic_symptom_vomitfg|aic_symptom_nausea|bleeding|impaired_mental_status|hepatomegaly|rash", names(R01_lab_results))], na.rm = TRUE)    #probable dengue
R01_lab_results$dengue_warning_signs<-rowSums(R01_lab_results[, grep("aic_symptom_impaired_mental_status|bleeding|aic_symptom_vomiting|aic_symptom_abdominal_pain|hepatomegaly|splenomegaly|edema", names(R01_lab_results))], na.rm = TRUE)#warning signs (dengue)
R01_lab_results$dengue_warning<-ifelse(R01_lab_results$probable_dengue >= 2 & R01_lab_results$dengue_warning_signs >0 , 1, 0)#dengue with warning signs   
table(R01_lab_results$dengue_warning_signs, R01_lab_results$probable_dengue)
# PCR -----------------------------------------------------------------
#PCR DENV or chikv
R01_lab_results$denv_chikv_result_pcr[R01_lab_results$result_pcr_denv_kenya == 0|R01_lab_results$result_pcr_denv_stfd == 0|R01_lab_results$result_pcr_chikv_kenya == 0|R01_lab_results$result_pcr_chikv_stfd == 0] <- 0
R01_lab_results$denv_chikv_result_pcr[R01_lab_results$result_pcr_denv_kenya == 1|R01_lab_results$result_pcr_denv_stfd == 1|R01_lab_results$result_pcr_chikv_kenya == 1|R01_lab_results$result_pcr_chikv_stfd == 1] <- 1
table(R01_lab_results$denv_chikv_result_pcr)

# UFI -----------------------------------------------------------------
# ufi1
R01_lab_results <- within(R01_lab_results, denv_result_ufi[R01_lab_results$ufi2_result_denv >45] <- 0)
R01_lab_results <- within(R01_lab_results, denv_result_ufi[R01_lab_results$ufi2_result_denv <=45] <- 1)
R01_lab_results <- within(R01_lab_results, denv_result_ufi[R01_lab_results$denv_result_ufi==1] <- 1)

R01_lab_results <- within(R01_lab_results, chikv_result_ufi[R01_lab_results$chikv_ct_ufi2 >45] <- 0)
R01_lab_results <- within(R01_lab_results, chikv_result_ufi[R01_lab_results$chikv_ct_ufi2 <=45] <- 1)
R01_lab_results <- within(R01_lab_results, chikv_result_ufi[R01_lab_results$chikv_result_ufi==1] <- 1)

# ufi2
R01_lab_results$ufi2_result_denv<-NA
R01_lab_results <- within(R01_lab_results, ufi2_result_denv[R01_lab_results$denv_ct_ufi2 >45] <- 0)
R01_lab_results <- within(R01_lab_results, ufi2_result_denv[R01_lab_results$denv_ct_ufi2 <=45] <- 1)

R01_lab_results$ufi2_result_chikv<-NA
R01_lab_results <- within(R01_lab_results, ufi2_result_chikv[R01_lab_results$chikv_ct_ufi2 >45] <- 0)
R01_lab_results <- within(R01_lab_results, ufi2_result_chikv[R01_lab_results$chikv_ct_ufi2 <=45] <- 1)

##age group---------------------------------------------------
R01_lab_results$age = R01_lab_results$age_calc_rc  # your new merged column starts with age_calc_rc
R01_lab_results$age[!is.na(R01_lab_results$aic_calculated_age)] = R01_lab_results$aic_calculated_age[!is.na(R01_lab_results$aic_calculated_age)]  # merge with aic_calculated_age
R01_lab_results$age[!is.na(R01_lab_results$age_calc)] = R01_lab_results$age_calc[!is.na(R01_lab_results$age_calc)]  # merge with age_calc

R01_lab_results$age<-round(R01_lab_results$age)

summary(R01_lab_results$age)
R01_lab_results$age_group<-NA
R01_lab_results <- within(R01_lab_results, age_group[age<=2] <- "under 2")
R01_lab_results <- within(R01_lab_results, age_group[age>2 & age<=5] <- "2-5")
R01_lab_results <- within(R01_lab_results, age_group[age>5 & age<=10] <- "6-10")
R01_lab_results <- within(R01_lab_results, age_group[age>10 & age<=15] <- "11-15")
R01_lab_results <- within(R01_lab_results, age_group[age>15] <- "over 15")
R01_lab_results$age_group <- factor(R01_lab_results$age_group, levels = c("under 2", "2-5", "6-10", "11-15", "over 15"))
table(R01_lab_results$age_group, exclude = NULL)

# deidentify -----------------------------------------------------------------
#take name out of event.
#names(R01_lab_results)[names(R01_lab_results) == 'redcap_event_name'] <- 'redcap_event'
#identifiers<-grep("name|gps", names(R01_lab_results), value = TRUE)
#R01_lab_results<-R01_lab_results[ , !(names(R01_lab_results) %in% identifiers)]#turn of the deidentifiers to export the u24 data. 
#names(R01_lab_results)[names(R01_lab_results) == 'redcap_event'] <- 'redcap_event_name'

# incidence -----------------------------------------------------------------
table(R01_lab_results$seroc_denv_kenya_igg, R01_lab_results$seroc_denv_stfd_igg, exclude=NULL)

R01_lab_results$id_cohort<-substr(R01_lab_results$person_id, 2, 2)
R01_lab_results$id_city<-substr(R01_lab_results$person_id, 1, 1)


#prnt
R01_lab_results <- within(R01_lab_results, prnt_80_chikv[R01_lab_results$prnt_80_chikv =="<10"] <- "5")
R01_lab_results <- within(R01_lab_results, prnt_80_wnv[R01_lab_results$prnt_80_wnv ==">80"] <- "160")
R01_lab_results <- within(R01_lab_results, prnt_80_wnv[R01_lab_results$prnt_80_wnv =="No sample"] <- NA)
R01_lab_results <- within(R01_lab_results, prnt_80_denv[R01_lab_results$prnt_80_denv =="<10"] <- "5")
R01_lab_results <- within(R01_lab_results, prnt_80_denv[R01_lab_results$prnt_80_denv =="nd"] <- "NA")
R01_lab_results <- within(R01_lab_results, prnt_80_wnv[R01_lab_results$prnt_80_wnv =="<10"] <- "5")
R01_lab_results <- within(R01_lab_results, prnt_80_onn[R01_lab_results$prnt_80_onn ==">80"] <- "160")
R01_lab_results <- within(R01_lab_results, prnt_80_onn[R01_lab_results$prnt_80_onn =="<10"] <- "5")
R01_lab_results <- within(R01_lab_results, prnt_80_onn[R01_lab_results$prnt_80_onn =="No sample"|R01_lab_results$prnt_80_onn =="no sample"] <- NA)
R01_lab_results$prnt_80_wnv<-as.numeric(as.character(R01_lab_results$prnt_80_wnv))

R01_lab_results$prnt_80_onn<-as.numeric(as.character(R01_lab_results$prnt_80_onn))
R01_lab_results$prnt_result_onn<-NA
R01_lab_results <- within(R01_lab_results, prnt_result_onn[R01_lab_results$prnt_80_onn <10] <- 0)
R01_lab_results <- within(R01_lab_results, prnt_result_onn[R01_lab_results$prnt_80_onn >=10] <- 1)


R01_lab_results$prnt_80_denv<-as.numeric(as.character(R01_lab_results$prnt_80_denv))
R01_lab_results$prnt_result_denv<-NA
R01_lab_results <- within(R01_lab_results, prnt_result_denv[R01_lab_results$prnt_80_denv <10] <- 0)
R01_lab_results <- within(R01_lab_results, prnt_result_denv[R01_lab_results$prnt_80_denv >=10] <- 1)
table(R01_lab_results$prnt_result_denv, R01_lab_results$prnt_80_denv)

R01_lab_results$prnt_80_chikv<-as.numeric(as.character(R01_lab_results$prnt_80_chikv))
class(R01_lab_results$prnt_80_chikv)
table(R01_lab_results$prnt_80_chikv)
R01_lab_results$prnt_result_chikv<-NA
R01_lab_results <- within(R01_lab_results, prnt_result_chikv[R01_lab_results$prnt_80_chikv <10] <- 0)
R01_lab_results <- within(R01_lab_results, prnt_result_chikv[R01_lab_results$prnt_80_chikv >=10] <- 1)
table(R01_lab_results$prnt_result_chikv)

R01_lab_results$prnt_80_wnv<-as.numeric(as.character(R01_lab_results$prnt_80_wnv))
class(R01_lab_results$prnt_80_wnv)
table(R01_lab_results$prnt_80_wnv)
R01_lab_results$prnt_result_wnv<-NA
R01_lab_results <- within(R01_lab_results, prnt_result_wnv[R01_lab_results$prnt_80_wnv <10] <- 0)
R01_lab_results <- within(R01_lab_results, prnt_result_wnv[R01_lab_results$prnt_80_wnv >=10] <- 1)
table(R01_lab_results$prnt_result_wnv)

#use tested = 1 as the zero for infection.
#stfd denv igg seroconverters or PCR or UFI/UFI2 positives as infected. 
R01_lab_results$infected_denv_stfd[R01_lab_results$tested_denv_stfd_igg ==1 |R01_lab_results$result_pcr_denv_kenya==0|R01_lab_results$result_pcr_denv_stfd==0|R01_lab_results$denv_result_ufi==0|R01_lab_results$prnt_result_denv==0|R01_lab_results$ufi2_result_denv==0]<-0
#should we excclude the igg only for coast? this infaltes our denominator, especially on the coast. 
R01_lab_results$infected_denv_stfd[R01_lab_results$seroc_denv_stfd_igg==1|R01_lab_results$result_pcr_denv_kenya==1|R01_lab_results$result_pcr_denv_stfd==1|R01_lab_results$denv_result_ufi==1|R01_lab_results$prnt_result_denv==1|R01_lab_results$ufi2_result_denv==1]<-1
table(R01_lab_results$infected_denv_stfd)


#stfd chikv igg seroconverters or PCR or UFI/UFI2 positives as infected. or PNRT +
R01_lab_results$infected_chikv_stfd[R01_lab_results$tested_chikv_stfd_igg ==1 |R01_lab_results$result_pcr_chikv_kenya==0|R01_lab_results$chikv_result_ufi==0|R01_lab_results$prnt_result_chikv==0|R01_lab_results$ufi2_result_chikv==0]<-0
R01_lab_results$infected_chikv_stfd[R01_lab_results$seroc_chikv_stfd_igg==1|R01_lab_results$result_pcr_chikv_kenya==1|R01_lab_results$chikv_result_ufi==1|R01_lab_results$prnt_result_chikv==1|R01_lab_results$ufi2_result_chikv==1]<-1
table(R01_lab_results$infected_chikv_stfd)

#chikv or denv incidence
R01_lab_results$infected_denv_chikv_stfd[R01_lab_results$infected_chikv_stfd==0 |R01_lab_results$infected_denv_stfd==0]<-0
R01_lab_results$infected_denv_chikv_stfd[R01_lab_results$infected_chikv_stfd==1 |R01_lab_results$infected_denv_stfd==1]<-1
table(R01_lab_results$infected_denv_chikv_stfd)

##denominator is only those tested for chikv by pcr or igg at stfd ---------------------------------------------------
R01_lab_results$tested_chikv<-NA
R01_lab_results<- within(R01_lab_results, tested_chikv[infected_chikv_stfd==1 |tested_chikv_stfd_igg==1 | !is.na(result_pcr_chikv_kenya)| !is.na(result_pcr_chikv_stfd)| !is.na(chikv_result_ufi)|!is.na(prnt_80_chikv)] <- 1)
table(R01_lab_results$infected_chikv_stfd, exclude = NULL)
table(R01_lab_results$tested_chikv, exclude = NULL)
table(R01_lab_results$tested_chikv, R01_lab_results$infected_chikv_stfd, exclude = NULL)
(19/3938)*100 #incidence chikv

##denominator is only those tested for denv by pcr or igg at stfd---------------------------------------------------
R01_lab_results$tested_denv<-NA
R01_lab_results<- within(R01_lab_results, tested_denv[infected_denv_stfd==1 |tested_denv_stfd_igg==1 | !is.na(result_pcr_denv_kenya) | !is.na(result_pcr_denv_stfd)|!is.na(denv_result_ufi)|!is.na(prnt_80_denv)] <- 1)
#denominator is only those tested for denv or chikv by pcr or igg at stfd
R01_lab_results$tested_denv_chikv<-NA
R01_lab_results<- within(R01_lab_results, tested_denv_chikv[tested_chikv==1 |tested_denv==1] <- 1)

# prevalence -----------------------------------------------------------------
R01_lab_results_prev_denv<-R01_lab_results[which(R01_lab_results$infected_denv_stfd==1|R01_lab_results$result_igg_denv_stfd==1),]
R01_lab_results_prev_denv <- R01_lab_results_prev_denv[order(R01_lab_results_prev_denv$person_id, R01_lab_results_prev_denv$int_date),]
R01_lab_results_prev_denv <- R01_lab_results_prev_denv[!duplicated(R01_lab_results_prev_denv$person_id),]
R01_lab_results_prev_denv$prev_denv_stfd<-"1"
R01_lab_results_prev_denv<-R01_lab_results_prev_denv[ , grepl( "person_id|redcap_event|prev_denv_stfd" , names(R01_lab_results_prev_denv) ) ]

R01_lab_results<-  merge(R01_lab_results_prev_denv, R01_lab_results, by = c("person_id","redcap_event_name"), all.y = TRUE)

R01_lab_results_prev_chikv<-R01_lab_results[which(R01_lab_results$infected_chikv_stfd==1|R01_lab_results$result_igg_chikv_stfd==1),]
R01_lab_results_prev_chikv <- R01_lab_results_prev_chikv[order(R01_lab_results_prev_chikv$person_id, R01_lab_results_prev_chikv$int_date),]
R01_lab_results_prev_chikv <- R01_lab_results_prev_chikv[!duplicated(R01_lab_results_prev_chikv$person_id),]
R01_lab_results_prev_chikv$prev_chikv_stfd<-1
R01_lab_results_prev_chikv<-R01_lab_results_prev_chikv[ , grepl( "person_id|redcap_event|prev_chikv_stfd" , names(R01_lab_results_prev_chikv) ) ]
R01_lab_results<-  merge(R01_lab_results_prev_chikv, R01_lab_results, by = c("person_id","redcap_event_name"), all.y = T)

R01_lab_results <- within(R01_lab_results, prev_chikv_stfd[(tested_chikv==1 & is.na(prev_chikv_stfd))|(!is.na(result_igg_chikv_stfd) & is.na(prev_chikv_stfd)) ] <- 0)
R01_lab_results <- within(R01_lab_results, prev_denv_stfd[(tested_denv==1 & is.na(prev_denv_stfd))|(!is.na(result_igg_denv_stfd) & is.na(prev_denv_stfd)) ] <- 0)

table(R01_lab_results$prev_chikv_stfd,R01_lab_results$id_cohort)
table(R01_lab_results$prev_denv_stfd, R01_lab_results$id_cohort)

# seroprevalence -----------------------------------------------------------------
table(R01_lab_results$result_igg_denv_stfd)
table(R01_lab_results$result_igg_chikv_stfd)

table(R01_lab_results$result_igg_chikv_stfd, R01_lab_results$site, exclude=NULL)
table(R01_lab_results$result_igg_denv_stfd, R01_lab_results$site)

table(R01_lab_results$result_igg_chikv_stfd, R01_lab_results$rural)
table(R01_lab_results$result_igg_denv_stfd, R01_lab_results$rural)

table(R01_lab_results$result_igg_denv_stfd, R01_lab_results$rural, exclude=NULL)

table(R01_lab_results$result_igg_chikv_stfd, R01_lab_results$id_cohort)
table(R01_lab_results$result_igg_denv_stfd, R01_lab_results$id_cohort)


#save file
save(R01_lab_results,file="R01_lab_results.rda")
R01_lab_results$gender_all = R01_lab_results$gender  # your new merged column start with gender
R01_lab_results$gender_all[!is.na(R01_lab_results$gender_aic)] = R01_lab_results$gender_aic[!is.na(R01_lab_results$gender_aic)]  # merge with gender_aic
table(R01_lab_results$gender_all, exclude = NULL)

# meds prescribed ---------------------------------------------------------
R01_lab_results$malaria_treatment_other_kenya<-gsub(",", "_", R01_lab_results$malaria_treatment_other_kenya)
R01_lab_results$all_meds<-NA
R01_lab_results$all_meds<-paste(R01_lab_results$meds_prescribed, R01_lab_results$malaria_treatment_other_kenya, sep="_")

R01_lab_results$all_meds<-tolower(R01_lab_results$all_meds)
R01_lab_results$all_meds <- gsub('  ', ' ', R01_lab_results$all_meds)
R01_lab_results$all_meds <- gsub(' ', '_', R01_lab_results$all_meds)

R01_lab_results$all_meds<-gsub("apirin|diclonac|voline_gel|voltaren|dinac|duclofenac|ibeufen|buscopan|painmedi|brufen|painmedi|painmedbrufen|diclo|ibrufen|ibrufeen", " painmed ", R01_lab_results$all_meds)
R01_lab_results$all_meds<-gsub("plasil|actals|actal", " gerd ", R01_lab_results$all_meds)
R01_lab_results$all_meds<-gsub("ventolin|ventoli|sabutanol|salbutamol|albutol|salbutanol|butanol", " anti_bronchospasm ", R01_lab_results$all_meds)
R01_lab_results$all_meds<-gsub("amoxicillin|t.flamox|doxycline|butol|augumrnting|antibacterialardrop|antibacterialflaxin|cefriaxone|antibacterialeardrop|antibacterialfloxaxilin|flaxin|floxaxilin|antibacterialg|amoxyl|amoxl|amixil|erytthromycin|doxy|antibacterialflaxin|aoxil|seotrin|amoxil|penicillin|antibacterialeardrop|antibacterialg|chloramphenicol|teo|t.e.o|flagyl|flaggyl|ciproxin|augumentin|cefxime|ciproxin|chlamphenicol|cefixime|ciproxin|ciprofloxin|intravenous_metronidazole|nitrofurantion|ciprofloxacin|flagyla|gentamicin|metronidazole|floxapen|flucloxacill|trinidazole|vedrox|ampiclox|cloxacillin|ampicillin|albendaxole|albedazole|tinidazole|tetracycline|augmentin|ceftriaxone|penicillian|septrin|antibiotic|ceftrizin|cotrimoxazole|cefuroxime|erythromycin|gentamycin|cipro|antibacterial|septin|amoxxil|antibacteriale", " antibacterial ", R01_lab_results$all_meds)
R01_lab_results$all_meds<-gsub("guaifenesin|expectants|expectant|tricoff|ossthial|osthial|expen", " expectorant ", R01_lab_results$all_meds)
R01_lab_results$all_meds<-gsub("syrup|unibrolcoldcap|unibrol|tricohist|trichohist|cold_cap|ascoril", " cough ", R01_lab_results$all_meds)
R01_lab_results$all_meds<-gsub( "cetrizine hydrochloride|chlorepheramine|chlore|hydrocrt|hydrocortisone|cetrizine|pirtion|piriton|priton|hydroctisone_cream|hydroctisone|hydroctione|cpm|pitriton|pirion|pirito|probeta-n|allergy|allergyllergy|hydrocort|promethazon|cezine|cetrizin|cezzzine|benahist|amine|adreline", " allergy ", R01_lab_results$all_meds)
R01_lab_results$all_meds<-gsub("calaminelotion|calamine|calaminetopical|calamine_lotion|cream|lotion|eye_ointment", " topical ", R01_lab_results$all_meds)
R01_lab_results$all_meds<-gsub("mulitivit|multivit|multivitamin|zinc_tablet|vitamin|multivit|multivt|zinc|multisupplement|supplement|m/supplements|ranferon|ferrous_sulphate|ferous|folic_acid|folic|ferrous|haemoton|saferon|mulitisupplement|raferon|multvit|mulit|mult|m/vits", " supplement ", R01_lab_results$all_meds)
R01_lab_results$all_meds<-gsub("paracentamol|paracetamol|ibuprofen|diclofeanc|diclofenac|calpol|plcetantipyretic", " antipyretic ", R01_lab_results$all_meds)
R01_lab_results$all_meds<-gsub("ketoconazole|griseofulvin|clotrimazole|clotrimazone|grisofluvin|graeofulvin|graseofulvin|greseofulvin|nystatin|nystatin_oral_mouth_paint|antifungaltopical|nestatin|ketaconazole|statin", " antifungal ", R01_lab_results$all_meds)
R01_lab_results$all_meds<-gsub("ors|o.r.s.", " ors ", R01_lab_results$all_meds)
R01_lab_results$all_meds<-gsub("i.v.|ivs|i.v.s.|i.v", " iv ", R01_lab_results$all_meds)
R01_lab_results$all_meds<-gsub("diloxanide", " antiamoeba ", R01_lab_results$all_meds)
R01_lab_results$all_meds<-gsub("admission|admitted|admit", " admit ", R01_lab_results$all_meds)
R01_lab_results$all_meds<-gsub("azt", " antiviral ", R01_lab_results$all_meds)
R01_lab_results$all_meds<-gsub("amitripin", " antidepressant ", R01_lab_results$all_meds)
R01_lab_results$all_meds<-gsub("predison", " immunosuppressant ", R01_lab_results$all_meds)
R01_lab_results$all_meds<-gsub("paraziquantel|albendazole|abz|mebendazole|benzimidazole|diloxanide|paraziquatel|fluconazole|bendazole", " antihelmenthic ", R01_lab_results$all_meds)
R01_lab_results$all_meds<-gsub("plcet|plct|pct|pcet|irin|coarem|plcet|im_quinine|attesunate|artesunate|artesun|quinine|coartem|coatem|antimalarial|caortem|coarterm|coart|coaterm|coartm|quinnie|atersunate|quinnine|paludrin|quinnie|duocotecxin|pheramine|artsun|atesunate|atesa|artesinate|pcm", " antimalarial ", R01_lab_results$all_meds)
#sp and al are too short to include.
R01_lab_results$all_meds<-gsub("/|none|other|and|eardrop", " ", R01_lab_results$all_meds)

# dummy meds vars ---------------------------------------------------------
#subset meds
meds<-R01_lab_results[, c("person_id", "redcap_event_name","all_meds", "meds_prescribed","malaria_treatment_other_kenya")]
meds <- lapply(meds, function(x) {  gsub("NA", "", x)})
meds <- lapply(meds, function(x) {  gsub("na", "", x)})
meds <- lapply(meds, function(x) {  gsub(" ", "_", x)})
meds <- lapply(meds, function(x) {  gsub(", ", "_", x)})
meds <- lapply(meds, function(x) {  gsub("__", "_", x)})
meds<-as.data.frame(meds)
#create dummy vars for all meds
lev <- levels(factor(meds$all_meds))
lev <- unique(unlist(strsplit(lev, "_")))
mnames <- gsub(" ", "_", paste("med", lev, sep = "_"))
result <- matrix(data = "0", nrow = length(meds$all_meds), ncol = length(lev))
char.med <- as.character(meds$all_meds)
for (i in 1:length(lev)) {
  result[grep(lev[i], char.med, fixed = TRUE), i] <- "1"
}
result <- data.frame(result, stringsAsFactors = TRUE)
colnames(result) <- mnames
meds <- cbind(meds,result)
R01_lab_results<-merge(meds,R01_lab_results,by=c("person_id","redcap_event_name"), all=TRUE)    

subset( R01_lab_results, med_cal==1, c(all_meds.x,meds_prescribed.x,  malaria_treatment_other_kenya.x,person_id))
R01_lab_results$malaria_treatment_other_kenya.x
R01_lab_results <- within(R01_lab_results, med_antimalarial[R01_lab_results$malaria_treatment_kenya___1==1] <- 1)
R01_lab_results <- within(R01_lab_results, med_antimalarial[R01_lab_results$malaria_treatment_kenya___2==1] <- 1)
R01_lab_results <- within(R01_lab_results, med_antimalarial[R01_lab_results$malaria_treatment_kenya___3==1] <- 1)
R01_lab_results <- within(R01_lab_results, med_antimalarial[R01_lab_results$malaria_treatment_kenya___4==1] <- 1)
R01_lab_results <- within(R01_lab_results, med_antimalarial[R01_lab_results$malaria_treatment_kenya___5==1] <- 1)
table(R01_lab_results$med_antimalarial)

R01_lab_results$antiparasite <- NA
R01_lab_results <- within(R01_lab_results, antiparasite[R01_lab_results$med_antihelmenthic==0|R01_lab_results$med_antimalarial==0] <- 0)
R01_lab_results <- within(R01_lab_results, antiparasite[R01_lab_results$med_antihelmenthic==1|R01_lab_results$med_antimalarial==1] <- 1)
table(R01_lab_results$antiparasite)

library(stringr)
R01_lab_results$number_meds <- str_count(R01_lab_results$all_meds.x, "_")
R01_lab_results$number_meds<-R01_lab_results$number_meds+1
table(R01_lab_results$number_meds)

# save cleaned dataset ----------------------------------------------------
setwd("C:/Users/amykr/Box Sync/Amy Krystosik's Files/Data Managment/redcap/ro1 lab results long")
save(R01_lab_results,file="C:/Users/amykr/Box Sync/Amy Krystosik's Files/david coinfection paper/data/R01_lab_results.david.coinfection.dataset.rda")    #save as r data frame for use in other analysis. 
```


#Define strata
```{r}
library(tidyr)
grep("denv", names(R01), value=TRUE)
#Find DENV PCR positives
table(R01$denv_result_ufi, R01$denv_result_ufi2, exclude = NULL)
str(R01$denv_result_ufi)
#denv_result_ufi is a logical

print(R01$denv_result_ufi[R01$denv_result_ufi=="1"])
str(R01)

```


# Used different selection criteria
# Used all variables-->exclude sparse data n<5, include only ones with p<=0.002
```{r}

percentsbackup<- percents 
#10 variables with between strata differences in frquency after excluding sparse data
vars10 <-c("body_ache", "aic_symptom_general_body_ache", "aic_symptom_sick_feeling", "aic_pe_lethargic", "aic_symptom_runny_nose", "aic_symptom_vomiting", "aic_symptom_abdominal_pain", "aic_pe_red", "aic_pe_tender", "aic_pe_cervial")
#Variables after excluding the ones with sparse data n<5 after stratification
#("body_ache", "aic_symptom_general_body_ache", "aic_symptom_sick_feeling", "aic_symptom_loss_of_appetite", "aic_symptom_chills", "aic_pe_lethargic", aic_symptom_fever, aic_symptom_headache, aic_symptom_dizziness, aic_pe_decreased_strength, aic_pe_red_eyes, "aic_symptom_runny_nose", aic_symptom_sore_throat, aic_symptom_cough, aic_pe_rapid_breathing, aic_pe_rapid_rate, aic_symptom_nausea, "aic_symptom_vomiting", "aic_symptom_abdominal_pain", aic_symptom_diarrhea, aic_symptom_joint_pains, "aic_pe_red", "aic_pe_tender", aic_pe_icteric_sclerae, aic_pe_splenomegaly, aic_pe_large_lymph_nodes, "aic_pe_cervial")  
```
# Adapting Amy's code
```{r}
#checking data
table(percents$strata_all, percents$malaria, exclude = NULL) #malaria captures any malaria positive
table(percents$strata_all, percents$infected_denv_stfd, exclude = NULL) #infected_denv_stfd captures any DENV infection
#percents<-percentsbackup #if need to reset database
library(epiR)
library(dplyr)
attach(percents)
#order factors ---------------------------------------------
#percents$strata_all <- factor(percents$strata_all, levels = c("malaria_pos_denv_pos", "malaria_neg_denv_pos", "malaria_pos_denv_neg","malaria_neg_denv_neg"))
#The line above results in all strata having NA!
percents$aic_symptom_dizziness <- factor(percents$aic_symptom_dizziness, levels = c(1, 0))
percents$aic_pe_decreased_strength <- factor(percents$aic_pe_decreased_strength, levels = c(1, 0))
percents$aic_pe_icteric_sclerae <- factor(percents$aic_pe_icteric_sclerae, levels = c(1, 0))
percents$aic_symptom_abdominal_pain <- factor(percents$aic_symptom_abdominal_pain, levels = c(1, 0))
percents$aic_pe_swollen <- factor(percents$aic_pe_swollen, levels = c(1, 0))
percents$aic_pe_tender <- factor(percents$aic_pe_tender, levels = c(1, 0))
percents$infected_denv_stfd <- factor(percents$infected_denv_stfd, levels = c(1, 0))

```

```{r}

prop.test(table(percents$strata_all, percents$aic_symptom_abdominal_pain), p = NULL,
          alternative = c("two.sided"),
          conf.level = 0.998, correct = TRUE)

prop.trend.test(table(percents$strata_all, percents$aic_symptom_abdominal_pain), p = NULL,
          alternative = c("two.sided"),
          conf.level = 0.95, correct = TRUE)

library(tableone)
CreateTableOne(vars=vars10, strata="strata_all", data=percents, factorVars=vars10, includeNA = FALSE, test = TRUE)
#,
  test = TRUE, testApprox = chisq.test, testExact = fisher.test, 
  smd = TRUE)
table(percents$strata_all)

percents % > %
	group_by(strata_all) % > %
	summarise(aic_symptom_abdominal_pain = mean(HR))



(tab2 <- prop.table(table(aic_symptom_abdominal_pain, strata_all), 1)*100)




#Logistic model with strata_all
test<-glm(aic_symptom_abdominal_pain~factor(strata_all), family="binomial", data = pedsql)
fits <- lapply(vars, function(x) {glm(substitute(i~factor(strata_all), list(i = as.name(x))), family="binomial", data = pedsql)})
exp(coefficients(test))
exp(confint(test))

coef<-lapply(fits, coefficients)
coef<-lapply(fits, function(x) {exp(cbind("Odds ratio" = coef(x), confint(x, level = 0.998, parallel="multicore", ncpus=4)))})

lapply(coef, function(x) write.table( data.frame(x), 'GLM-OR-998.csv'  , append= T, sep=',' ))
#GLM interaction between malaria and denv
effectmod<-glm(aic_symptom_abdominal_pain~malaria+infected_denv_stfd+malaria*infected_denv_stfd, family="binomial", data = pedsql)
effectmod <- lapply(vars, function(x) {glm(substitute(i~malaria+infected_denv_stfd+malaria*infected_denv_stfd, list(i = as.name(x))), family="binomial", data = pedsql)})
exp(coefficients(effectmod))
exp(confint(effectmod))
coef<-lapply(effectmod, coefficients)
coef<-lapply(effectmod, function(x) {exp(cbind("Odds ratio" = coef(x), confint(x, level = 0.998, parallel="multicore", ncpus=4)))})
lapply(coef, function(x) write.table( data.frame(x), 'GLM-OR-998-effectmod.csv'  , append= T, sep=',' ))


effectmod<-glm(aic_symptom_dizziness~malaria+infected_denv_stfd+malaria*infected_denv_stfd, family="binomial", data = pedsql)
exp(coefficients(effectmod))
exp(confint(effectmod))
```




#GLM
```{r}
#Logistic model with strata_all
test<-glm(aic_symptom_abdominal_pain~factor(strata_all), family="binomial", data = pedsql)
fits <- lapply(vars, function(x) {glm(substitute(i~factor(strata_all), list(i = as.name(x))), family="binomial", data = pedsql)})
exp(coefficients(test))
exp(confint(test))

coef<-lapply(fits, coefficients)
coef<-lapply(fits, function(x) {exp(cbind("Odds ratio" = coef(x), confint(x, level = 0.998, parallel="multicore", ncpus=4)))})

lapply(coef, function(x) write.table( data.frame(x), 'GLM-OR-998.csv'  , append= T, sep=',' ))
#GLM interaction between malaria and denv
effectmod<-glm(aic_symptom_abdominal_pain~malaria+infected_denv_stfd+malaria*infected_denv_stfd, family="binomial", data = pedsql)
effectmod <- lapply(vars, function(x) {glm(substitute(i~malaria+infected_denv_stfd+malaria*infected_denv_stfd, list(i = as.name(x))), family="binomial", data = pedsql)})
exp(coefficients(effectmod))
exp(confint(effectmod))
coef<-lapply(effectmod, coefficients)
coef<-lapply(effectmod, function(x) {exp(cbind("Odds ratio" = coef(x), confint(x, level = 0.998, parallel="multicore", ncpus=4)))})
lapply(coef, function(x) write.table( data.frame(x), 'GLM-OR-998-effectmod.csv'  , append= T, sep=',' ))


effectmod<-glm(aic_symptom_dizziness~malaria+infected_denv_stfd+malaria*infected_denv_stfd, family="binomial", data = pedsql)
exp(coefficients(effectmod))
exp(confint(effectmod))
```


# CMH for DENV affecting malaria by clinical variable
```{r}
 #CMH for all significant symptoms using alpha = 0.002
#Table 4---------------------------------------------

#install.packages("questionr")
library(questionr)
library(DescTools)
attach(pedsql)
#vars <-c("body_ache", "aic_symptom_general_body_ache", "aic_symptom_sick_feeling", "aic_pe_lethargic", "aic_symptom_runny_nose", "aic_symptom_vomiting", "aic_symptom_abdominal_pain", "aic_pe_red", "aic_pe_tender", "aic_pe_cervial")

#bodyache
odds.ratio(table(malaria,body_ache), level =.998)
  apply(table(malaria,body_ache,infected_denv_stfd), 3, odds.ratio,level=.998)
  mantelhaen.test(table(malaria,body_ache,infected_denv_stfd),correct = TRUE, conf.level = 0.998)  
  BreslowDayTest(table(malaria,body_ache,infected_denv_stfd), OR=NA, correct=TRUE)
#general body ache
odds.ratio(table(malaria,aic_symptom_general_body_ache), level =.998)
  apply(table(malaria,aic_symptom_general_body_ache,infected_denv_stfd), 3, odds.ratio,level=.998)
  mantelhaen.test(table(malaria,aic_symptom_general_body_ache,infected_denv_stfd),correct = TRUE,conf.level = 0.998)
    BreslowDayTest(table(malaria,aic_symptom_general_body_ache,infected_denv_stfd), OR=NA, correct=TRUE)
#sick feeling
odds.ratio(table(malaria,aic_symptom_sick_feeling), level =.998)
  apply(table(malaria,aic_symptom_sick_feeling,infected_denv_stfd), 3, odds.ratio,level=.998)
  mantelhaen.test(table(malaria,aic_symptom_sick_feeling,infected_denv_stfd),correct = TRUE,conf.level = 0.998)  
    BreslowDayTest(table(malaria,aic_symptom_sick_feeling,infected_denv_stfd), OR=NA, correct=TRUE)
#lethargy
odds.ratio(table(malaria,aic_pe_lethargic), level =.998)
  apply(table(malaria,aic_pe_lethargic,infected_denv_stfd), 3, odds.ratio,level=.998)
  mantelhaen.test(table(malaria,aic_pe_lethargic,infected_denv_stfd),correct=TRUE, conf.level = 0.998)
    BreslowDayTest(table(malaria,aic_pe_lethargic,infected_denv_stfd), OR=NA, correct=TRUE)
#Runny nose
odds.ratio(table(malaria,aic_symptom_runny_nose), level =0.998)
  apply(table(malaria,aic_symptom_runny_nose,infected_denv_stfd), 3, odds.ratio,level=0.998)
  mantelhaen.test(table(malaria,aic_symptom_runny_nose,infected_denv_stfd),correct=TRUE,conf.level = 0.998)  
    BreslowDayTest(table(malaria,aic_symptom_runny_nose,infected_denv_stfd), OR=NA, correct=TRUE)
   #vomiting
odds.ratio(table(malaria,aic_symptom_vomiting), level =0.998)
  apply(table(malaria,aic_symptom_vomiting,infected_denv_stfd), 3, odds.ratio,level=0.998)
  mantelhaen.test(table(malaria,aic_symptom_vomiting,infected_denv_stfd),correct=TRUE,conf.level = 0.998)
    BreslowDayTest(table(malaria,aic_symptom_vomiting,infected_denv_stfd), OR=NA, correct=TRUE)
  #abdominal pain
odds.ratio(table(malaria,aic_symptom_abdominal_pain), level =0.998)
  apply(table(malaria,aic_symptom_abdominal_pain,infected_denv_stfd), 3, odds.ratio,level=0.998)
  mantelhaen.test(table(malaria,aic_symptom_abdominal_pain,infected_denv_stfd),correct=TRUE,conf.level = 0.998)
    BreslowDayTest(table(malaria,aic_symptom_abdominal_pain,infected_denv_stfd), OR=NA, correct=TRUE)
 
    #joint erythema
  odds.ratio(table(malaria,aic_pe_red), level =0.998)
  apply(table(malaria,aic_pe_red,infected_denv_stfd), 3, odds.ratio,level=0.998)
  mantelhaen.test(table(malaria,aic_pe_red,infected_denv_stfd),correct=TRUE,conf.level = 0.998)
      BreslowDayTest(table(malaria,aic_pe_red,infected_denv_stfd), OR=NA, correct=TRUE)
      #joint tenderness --redo starting here
  odds.ratio(table(malaria,aic_pe_tender), level =0.998)
  apply(table(malaria,aic_pe_tender,infected_denv_stfd), 3, odds.ratio,level=0.998)
  mantelhaen.test(table(malaria,aic_pe_tender,infected_denv_stfd),correct=TRUE,conf.level = 0.998)
      BreslowDayTest(table(malaria,aic_pe_tender,infected_denv_stfd), OR=NA, correct=TRUE)
        #cervical adenopathy
  odds.ratio(table(malaria,aic_pe_cervial), level =0.998)
  apply(table(malaria,aic_pe_cervial,infected_denv_stfd), 3, odds.ratio,level=0.998)
  mantelhaen.test(table(malaria,aic_pe_cervial,infected_denv_stfd),correct=TRUE,conf.level = 0.998)
      BreslowDayTest(table(malaria,aic_pe_cervial,infected_denv_stfd), OR=NA, correct=TRUE)
        
```

```{r}
#BreslowDay
library(epiR)
library(DescTools)
attach(pedsql)
BreslowDayTest(table(malaria,body_ache,infected_denv_stfd), OR=NA, correct=FALSE)








if(!require(dplyr)){install.packages("dplyr")}
if(!require(DescTools)){install.packages("DescTools")}
if(!require(ggplot2)){install.packages("ggplot2")}
if(!require(grid)){install.packages("grid")}
if(!require(vcd)){install.packages("vcd")}

```
#GLMs
```{r}
GLmodel<-function(x) {glm(x ~ strata_all, family = binomial, data = pedsql)}
lapply(vars, GLmodel)
vars
exp(coef(y))
```

```{r}
#Table S4
#OR co-infect using alpha=0.002





#Table 2, OR of symptom/sign in reference to co-infection 
#List of 29 variables with significant chi square
variables <-c("body_ache", "aic_symptom_chills", "aic_symptom_headache", "aic_symptom_dizziness", "aic_pe_increased_reflexes", "aic_pe_decreased_strength", "aic_symptom_sore_throat", "aic_symptom_runny_nose", "aic_pe_nose_discharge", "aic_pe_eye_discharge", "aic_symptom_red_eyes", "aic_pe_icteric_sclerae", "aic_symptom_cough", "aic_pe_rales", "aic_pe_rapid_rate", "aic_symptom_nausea", "aic_symptom_vomiting", "aic_symptom_abdominal_pain", "aic_symptom_diarrhea","aic_symptom_rash", "aic_pe_localized_rash", "aic_pe_papular", "aic_symptom_joint_pains", "aic_pe_swollen", "aic_pe_red", "aic_pe_tender", "aic_pe_cervial", "aic_pe_large_lymph_nodes", "aic_pe_axillary")

#AIC <- fastDummies::dummy_cols(AIC, select_columns = "strata_all")
table(pedsql$strata_all)
#Logistic model with strata_all
test<-glm(aic_symptom_abdominal_pain~factor(strata_all), family="binomial", data = pedsql)
fits <- lapply(vars, function(x) {glm(substitute(i~factor(strata_all), list(i = as.name(x))), family="binomial", data = pedsql)})
exp(coefficients(test))
exp(confint(test))

coef<-lapply(fits, coefficients)
coef<-lapply(fits, function(x) {exp(cbind("Odds ratio" = coef(x), confint(x, level = 0.998, parallel="multicore", ncpus=4)))})

lapply(coef, function(x) write.table( data.frame(x), 'GLM-OR-998.csv'  , append= T, sep=',' ))

#https://stats.stackexchange.com/questions/63222/getting-p-values-for-multinom-in-r-nnet-package
#install.packages("afex")
library(afex)
set_sum_contrasts() # use sum coding, necessary to make type III LR tests valid
library(car)
#install.packages("AER")
library(AER)
p<-lapply(fits, coeftest)
library(broom)
ptable<-lapply(p, tidy)
lapply(ptable, function(x) write.table( data.frame(x), '191108_GLM_998.csv'  , append= T, sep=',' ))
```

```{r}
#save table of results
library(broom)
ptable<-lapply(p, tidy)
lapply(ptable, function(x) write.table( data.frame(x), 'pe_acute_OR_dec5.csv'  , append= T, sep=',' ))
```

